Deadlock Freedom on Self-Assembled DNA Nano Networks

### 2014-02-12 trunk/PAPER/implementation.tex
implementation updated

\item I/O Buffers-Trasceivers: These elements, one for
each port,  are responsible of data transmission between neighbours
nodes. The packets received or sent are stored in specifieds buffers
named input and output buffers respectivelly.

\item Switch Matrix: driven by a switch controller, enables the
reciprocal connections beetween the devices inside the node. Before
the creation of segments, such controller receives information from
DiSR block.  After this phase, the switch controller will be driven by
the block wich implements the routing algorithm. The Switch matrix is
essentially a series of multiplexers and demultiplexers.

\item : Processing Element (PE):Is strictly related to the node
functionality  and role inside a given network, e.g. being a
computation or storage node.

\item DiSR-Routing: The DiSR block contains all the hardware, control
logic and configuration registers, related to the implementation of the
proposed approach. The routing algorithm receives information
from DiSR  wich indicates the status of the segments that interests a
particular node. The routing operation will take in account of these
informations in order to perform a deadlock-free routing. Both Routing
and DiSR are connected to the switch matrix in order to receive
packets from the input buffer. Since the packets are processed one at
a time, a specific arbiter should be present within the switch matrix
controller.

@@

As discussed above, one of the main design challenge of DNA Self-
Assembled systems is the limited resource available to implement both
computations and routing decisions for each node. Assuming a budget of
$10^4$ CNFETs for each network's node~\cite{liu_jetcs}  we estimated
the required resources for implementing the entire DiSR block. The RTL
description of the circuitry described in the last section (with an
hardware description language)  of the DiSR circuitry has been written
and synthesized at gate-level. Considering the specific layout of each
single logic elements (NAND, full-adder, latch etc.), it has been
possible to get a rough estimate of the number of transistors
necessary to implement the DiSR logic. Figure~\ref{fig:imp_trend}
shows the results of synthesis in terms of number of devices (CNFETs)
versus the number of network nodes while the network scales up from
$10\times10$ to $100\times100$ nodes.

\begin{figure}
  \centering
  \includegraphics[width=0.50\textwidth]{pictures/synthesis.eps}
  \caption{Number of network's node (N) vs number of CNFETs.}
 \label{fig:synthesys}
\end{figure}

In particular, observing the Figure~\ref{fig:imp_trend} can 
be pointed out that the proposed implementation occupies about 17,5\% of the node budget. In particular the main contribution is due to the controller circuitry followed by LED registers.
Further, more than the absolute number of devices itself, it is interesting to
observe that the circuitry
complexity which implements the DiSR algorithm increases with a slowly
growing trend.  
An intuitive explanation of this is the relatively simple logic of
DiSR, whose behaviour is almost code on scalable storage structures.
For example, the number of registers
implementing the \emph{link\_visited[]} and \emph{link\_tvisited[]}
table follow the logarithmic function $N_{reg}=N_{port} \cdot
log_2(N)$ where \emph{Nport} is the number of the router’s ports, N is
the number of network’s nodes.

In order ha have estimate the maximum working clock frequency for the implemented
circuitry, the \cite{deng_isscc07}

>>>>

\item \textbf{I/O Buffers-Trasceivers:} These elements, one for
each port,  are responsible of data transmission between neighbours
nodes. The packets received or sent are stored in specifieds buffers
named input and output buffers respectivelly.

\item \textbf{Switch Matrix:} Driven by a switch controller, enables the
reciprocal connections beetween the devices inside the node. Before
the creation of segments, such controller receives information from
DiSR block.  After this phase, the switch controller will be driven by
the block wich implements the routing algorithm. The Switch matrix is
essentially a series of multiplexers and demultiplexers.

\item  \textbf{Processing Element (PE):} Is strictly related to the node
functionality  and role inside a given network, e.g. being a
computation or storage node.

\item \textbf{DiSR-Routing:} The DiSR block contains all the hardware, control
logic and configuration registers, related to the implementation of the
proposed approach. The routing algorithm receives information
from DiSR  wich indicates the status of the segments that interests a
particular node. The routing operation will take in account of these
informations in order to perform a deadlock-free routing. Both Routing
and DiSR are connected to the switch matrix in order to receive
packets from the input buffer. Since the packets are processed one at
a time, a specific arbiter should be present within the switch matrix
controller.

@@

As discussed above, one of the main design challenge of DNA Self-
Assembled systems is the limited resource available to implement both
computations and routing decisions for each node. Assuming a budget of
$10^4$ CNFETs for each network's node~\cite{liu_jetcs}  we estimated
the required resources for implementing the entire DiSR block. The RTL
description of the circuitry described in the last section (with an
hardware description language)  of the DiSR circuitry has been written
and synthesized at gate-level using \emph{Synopsys Design Compiler} with 
a generic technology library (GTECH).
Considering the specific layout of each single logic elements  (NAND,
full-adder, latch etc.), it has been possible to get a rough estimate
of the number of transistors necessary to implement the DiSR logic.
Figure~\ref{fig:synthesys} shows the results of synthesis in terms of
number of devices (CNFETs) versus the number of network nodes while
the network scales up from $10\times10$ to $100\times100$ nodes.

\begin{figure}
  \centering
  \includegraphics[width=0.50\textwidth]{pictures/synthesis.eps}
  \caption{Number of network's node (N) vs number of CNFETs necassary to implement
  the DiSR circuitry.}
 \label{fig:synthesys}
\end{figure}

In particular, observing the Figure~\ref{fig:synthesys} can  be
pointed out that the proposed implementation occupies about 17,5\% of
the node budget. In particular the main contribution is due to the
controller circuitry followed by LED registers. Further, more than the
absolute number of devices itself, it is interesting to observe that
the circuitry complexity which implements the DiSR algorithm increases
with a slowly growing trend.   An intuitive explanation of this is the
relatively simple logic of DiSR, whose behaviour is almost code on
scalable storage structures. For example, the number of registers
implementing the \emph{link\_visited[]} and \emph{link\_tvisited[]}
table follow the logarithmic function $N_{reg}=N_{port} \cdot
log_2(N)$ where \emph{Nport} is the number of the router’s ports, N is
the number of network’s nodes.

In order to have an idea of the maximum working clock frequency for the
implemented circuitry, we can obtain the timing results from a gate-
level netlist. Since a complete technology library usefull for 
commercial synthesis tool (like Design Compiler) doesn't exist for the this particular technology
(at the best that we know), we have synthesized the RTL description of DiSR
with a standard 32nm CMOS library obtain 
a delay of about $\mathrm{\tau_{DiSR}=10} FO4$ (fan-out of four).
Considering the results obtained in~\cite{deng_isscc07}, wich reports the
ratio in therms of FO4 between a standard 32nm CMOS technology and a
carbon nanotube ones, we can have a rough delay estimation.
In particolar has been reported a ratio R of: 
\begin{equation}
\mathrm{R=\frac{FO4_{CMOS}}{FO4_{CNFET}}=2} 
\end{equation}
Considering that F04 CMOS is about 30 ps, and the results of  
last equation thedelay of one FO4 for the CNFET case is about of
15 ps an then the working frequency can be cmputed as:

\begin{equation}
\mathrm{f_{clk}=\frac{1}{T_{max}}=\frac{1}{\tau_{DiSR} \cdot FO4_{CNFET}}}=
\frac{1}{10\cdot 15 \cdot 10^{-12}}=6,6 GHz
\end{equation}

where $\mathrm{\tau_{DiSR}}$ was expressed in therms of FO4.

Regards the power
consumption of the entire DiSR devices, we can skip to perform an a
